Paper 3 - Tables

### 2023-03-23 docs/causal_roadmap.qmd
Kept adding details to roadmap for triangulation. Started modifying files for tables and manuscript to use Quarto things

We will estimate the short-term effects of childhood exposure to non-persistent \acr{edcs} on the outcomes of interest measured in childhood. We will perform an ExWAS kind of analysis with the dependent variable $y$ being the outcome, and the independent variable $x_i$ being the levels of \acr{edc} $i$. We will assume that all associations are confounded by the same set of confounders, independently of the \acr{edc} considered.

@@

We will estimate the short-term effects of childhood exposure to non-persistent \acr{edcs} on metabolites and proteins measured in childhood. We will perform an XWAS analysis with the dependent variable $m_i$ being the metabolite or protein $i$, and the independent variable $x_j$ being the levels of \acr{edc} $j$. We will assume that all associations are confounded by the same set of confounders, independently of the \acr{edc} considered.

@@

    - We will use MR-Egger [@BowdenDaveySmithBurgess:2015], weighted median [@BowdenSmithHaycock:2016], and MR-PRESSO [@VerbanckChenNeale:2018] for sensitivity analyses.

    We will make use of \acr{2smr} [@DaveySmithEbrahim:2003; @PierceBurgess:2013; @DaveySmithHemani:2014]. We will thus employ results from two different types of \acr{gwass}: a exposure \acr{gwas}, using The Metabolomics GWAS Server [@SuhreShinPetersen:2011; @ShinFaumanPetersen:2014], and a outcome \acr{gwas}, using the OpenGWAS project [@10.7554/eLife.34408; @ElsworthLyonAlexander:2020]. The effects of the selected genetic variants will then be harmonized.

@@

@@

```{r}
source("../DAGs/dag_v1.R")
source("../code/dags.R")

dags <- dag_v1()
```

@@

```{r}
#| fig-cap: DAG for research question 1.
#| fig-align: center
#| fig-width: 30
#| fig-height: 40
#| out-extra: "keepaspectratio=true"
dag_rq <- dags$chem_to_out %>%
  dagitty::dagitty()
dag_rq %>%
  ggdag::tidy_dagitty() %>%
  ggdag::ggdag_status(text_size = 12, text_col = "black") +
  ggdag::scale_adjusted() +
  ggdag::theme_dag() +
  ggplot2::theme(legend.text = ggplot2::element_text(size = 20))

@@

```{r}
#| fig-cap: DAG for research question 2.
#| fig-align: center
#| fig-width: 30
#| fig-height: 40
#| out-extra: "keepaspectratio=true"
dag_rq <- dags$chem_to_out %>%
  dagitty::dagitty()
dag_rq %>%
  ggdag::tidy_dagitty() %>%
  ggdag::ggdag_status(text_size = 12, text_col = "black") +
  ggdag::scale_adjusted() +
  ggdag::theme_dag() +
  ggplot2::theme(legend.text = ggplot2::element_text(size = 20))

@@

@@

```{r}
#| fig-cap: DAG for research question 3.
#| fig-align: center
#| fig-width: 30
#| fig-height: 40
#| out-extra: "keepaspectratio=true"
dag_rq <- dags$chem_to_out %>%
  dagitty::dagitty()
dag_rq %>%
  ggdag::tidy_dagitty() %>%
  ggdag::ggdag_status(text_size = 12, text_col = "black") +
  ggdag::scale_adjusted() +
  ggdag::theme_dag() +
  ggplot2::theme(legend.text = ggplot2::element_text(size = 20))

>>>>

We will estimate the short-term effects of childhood exposure to non-persistent \acr{edcs} on the outcomes of interest measured in childhood. We will perform an ExWAS kind of analysis with the dependent variable $y$ being the outcome, and the independent variable $x_i$ being the levels of \acr{edc} $i$. We will assume that all associations are confounded by the same set of confounders, independently of the \acr{edc} considered. We will further stratify the analyses by sex, to account for potential sex-specific effects of the chemicals.

@@

We will estimate the short-term effects of childhood exposure to non-persistent \acr{edcs} on metabolites and proteins measured in childhood. We will perform an XWAS analysis with the dependent variable $m_i$ being the metabolite or protein $i$, and the independent variable $x_j$ being the levels of \acr{edc} $j$. We will assume that all associations are confounded by the same set of confounders, independently of the \acr{edc} considered. We will further stratify the analyses by sex, to account for potential sex-specific effects of the chemicals.

@@

    - We will use MR-Egger [@BowdenDaveySmithBurgess:2015], weighted median [@BowdenSmithHaycock:2016], and MR-PRESSO [@VerbanckChenNeale:2018] for sensitivity analyses. For *secondary* sensitivity analyses, we will use: Steiger filtering, Cochran's Q-statistics (to check for heterogeneity), and the Egger intercept (to check for horizontal pleiotropy) \textcolor{red}{(citations needed)}.

    We will make use of \acr{2smr} [@DaveySmithEbrahim:2003; @PierceBurgess:2013; @DaveySmithHemani:2014]. We will thus employ results from two different types of \acr{gwass}: a exposure \acr{gwas}, using The Metabolomics GWAS Server [@SuhreShinPetersen:2011; @ShinFaumanPetersen:2014], and a outcome \acr{gwas}, using the OpenGWAS project [@10.7554/eLife.34408; @ElsworthLyonAlexander:2020]. The effects of the selected genetic variants will then be harmonized. We will fit simple regression models to examine whether the selected \acr{ivs} predict the relevant nodes of the \acr{dag} (i.e., exposures and outcomes), using the F-statistics. Furthermore, we will fit simple regression models of the *form* `confounder ~ SNP`, adjusted for the *top* genetic \acr{pcs}, to test whether the \acr{snps} are separately associated with the confounders. As done in previous works, we will standardized all the considered outcomes. The results (both of the main and the sensitivity analyses) will be presented by means of forest plots, for the *significant* associations.

    In the case of \acr{1smr}, we will make use of the following criteria to select the \acr{snps} for both the exposures and the outcomes:
      - We will favour *cis*-variants.
      - We will *a priori* define a p-value threshold.
      - We will perform \acr{ld} clumping of the genetic variants.
      - We will *a priori* define a \acr{maf} threshold.
    Similarly, the models will be adjusted for the *top* genetic \acr{pcs}.

@@

\textcolor{red}{Is it possible to include models from the `brms` R package?}

@@

```{r}
source("../DAGs/dag_v2.R")

dags <- dag_v2()
```

@@

```{r}
#| fig-cap: DAG for research question 1.
#| fig-align: center
#| fig-width: 30
#| fig-height: 40
#| out-extra: "keepaspectratio=true"
dag_rq <- dags$chem_to_out

myphd::visualize_dag(dag_rq)
# dag_rq %>%
#   ggdag::tidy_dagitty() %>%
#   ggdag::node_status() %>%
#   ggplot2::ggplot(ggplot2::aes(x, y, 
#                                xend = xend, yend = yend, 
#                                color = status)) +
#   ggdag::geom_dag_edges() +
#   ggdag::geom_dag_point() +
#   geom_dag_text_repel() +
#   ggokabeito::scale_color_okabe_ito(na.value = "grey90") +
#   theme_dag() +
#   ggplot2::coord_cartesian(clip = "off")

@@

```{r}
#| fig-cap: DAG for research question 2.
#| fig-align: center
#| fig-width: 30
#| fig-height: 40
#| out-extra: "keepaspectratio=true"
```

@@

```{r}
```

@@

```{r}
#| fig-cap: DAG for research question 3.
#| fig-align: center
#| fig-width: 30
#| fig-height: 40
#| out-extra: "keepaspectratio=true"
```

### 2023-03-23 output/paper/tables.qmd
Kept adding details to roadmap for triangulation. Started modifying files for tables and manuscript to use Quarto things

---
title: "Tables"
output:
  bookdown::pdf_document2
---

```{r}
knitr::opts_chunk$set(echo = FALSE, 
                      warning = FALSE, message = FALSE, 
                      cache = TRUE)
```

<!-- ####################################################################### -->
<!-- Study populations -->
```{r}

```
<!-- ####################################################################### -->

<!-- ####################################################################### -->
<!-- Exposures description -->
```{r}

```
<!-- ####################################################################### -->

>>>>

---
title: "Paper 3 - Tables"
format:
  docx:
    reference-doc: reference_document.docx
  pdf:
    pdf-engine: pdflatex
    number-sections: true
    toc: false
    margin-left: 15mm
    margin-right: 15mm
    margin-bottom: 20mm
    margin-top: 15mm
    include-in-header:
      - \usepackage{rotating}
      - \usepackage{float}
      - \usepackage{typearea}
link-citations: true
linkcolor: blue
execute:
  echo: false
  warning: false
  error: false
  message: false
editor_options: 
  chunk_output_type: console
project:
  execute_dir: project
---

```{r}
#| include: false
here::i_am("tables.qmd")
source("../../../../all_phd/edcs.R")
library(tidyverse)
```

<!-- ####################################################################### -->
# Study populations
<!-- ####################################################################### -->
```{r}

```

<!-- ####################################################################### -->
# Exposures
<!-- ####################################################################### -->
```{r}
#| label: tbl-chem-info
#| tbl-cap: Information about the non-persistent EDCs.
edcs_info <- edcs_information() %>%
  dplyr::select(-c(chem_id, smiles))
gt::gt(edcs_info, groupname_col = "class") %>%
  gt::cols_label(
    full_name = "Compound", 
    short_name = "Symbol", 
    #smiles = "SMILES", 
    pubchem_cid = "PubChem CID", 
    ctd_id = "CTD ID", 
    exposome_explorer_id = "Exposome Explore ID", 
    parental_compound = "Parental compound"
  ) %>%
  gt::tab_style(
    locations = gt::cells_column_labels(columns = everything()), 
    style = list(
      gt::cell_borders(sides = "bottom", weight = gt::px(3)), 
      gt::cell_text(weight = "bold")
    )
  ) %>%
  gt::tab_style(
    locations = gt::cells_row_groups(group = everything()), 
    style = list(gt::cell_text(weight = "bold"))
  )
```
